---
title: "test"
author: "Ana Martins"
date: "27/03/2023"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r load libraries, echo=FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(directlabels)
library(psych)
library(brms)
```

# Load data, as.factor 'country' and inspect
```{r clean consumption data, echo=FALSE, message=FALSE, eval=FALSE}
data <- read.csv("test1.csv") %>% 
  mutate(country = as.factor(country))
glimpse(data)
```

# Group, filter and summarise data
```{r}
data.new <- as.data.frame(apply(data,
                                2,
                                function(x) gsub("\\s+","",x)))
data.new %>% 
  group_by(data.new$country) %>% 
  filter(country!="All" & country!="All ") %>% 
  summarise(number_of_papers = n_distinct(id),
            iucn_region = first(iucn_region))
```

# Rank countries and IUCN regions with more studies
```{r}
top10_country <- data.new %>% 
  group_by(country) %>%
  count() %>%
  arrange(desc(n))

top10_region <- data.new %>% 
  group_by(iucn_region) %>%
  count()%>%
  arrange(desc(n))
```

# Merge positive and negative columns
```{r}
correlates <- data.new %>%
  mutate(culinary_traditions = case_match(culinary_traditions_negative,
                                        "1" ~ "-1", 
                                        .default = culinary_traditions_positive),
         religious_traditions = case_match(religious_traditions_negative,
                                          "1" ~ "-1", 
                                        .default = religious_traditions_positive),
         sensory_perception = case_match(sensory_perception_negative,
                                         "1" ~ "-1", 
                                        .default = sensory_perception_positive),
         health_beliefs = case_match(health_beliefs_negative,
                                     "1" ~ "-1", 
                                        .default = health_beliefs_positive),
         eating_habits = case_match(eating_habits_negative,
                                    "1" ~ "-1", 
                                        .default = eating_habits_positive),
         preparation = case_match(preparation_negative,
                                  "1" ~ "-1", 
                                        .default = preparation_positive),
         self_efficacy = case_match(self_efficacy_negative,
                                    "1" ~ "-1", 
                                        .default = self_efficacy_positive),
         price_perception = case_match(price_perception_negative,
                                        "1" ~ "-1", 
                                        .default = price_perception_positive),
         fish_availability = case_match(fish_availability_negative,
                                         "1" ~ "-1", 
                                        .default = fish_availability_positive),
         marketing = case_match(marketing_negative,
                                 "1" ~ "-1", 
                                        .default = marketing_postive),
         livelihood = case_match(livelihood_negative,
                                 "1" ~ "-1", 
                                        .default = livelihood_positive))
```

# Select merged columns and make correlates numeric
```{r}
df <- correlates %>%
  select(iucn_region,
         country,
         iso3_code,
         mislabelling:livelihood)%>%
  mutate(across(mislabelling:livelihood, as.numeric))
str(df)
```

# Taking the mean of each correlate
```{r}
data.mean <- df %>% 
  group_by(country) %>%
  summarise(iucn_region = first(iucn_region),
            iso3_code = first(iso3_code),
            mislabelling = mean (mislabelling),
            culinary_traditions = mean (culinary_traditions),
            religious_traditions = mean (religious_traditions),
            sensory_perception = mean(sensory_perception),
            health_beliefs = mean(health_beliefs),
            eating_habits = mean(eating_habits),
            preparation = mean(preparation),
            self_efficacy = mean(self_efficacy),
            price_perception = mean (price_perception), 
            fish_availability = mean(fish_availability), 
            marketing = mean (marketing), 
            livelihood = mean (livelihood))
```

# Summarising columns and scaling correlates
```{r}
poo <- data.mean %>% 
  select(-c(iucn_region,iso3_code))%>%
  mutate(across(mislabelling:livelihood, ~scale(., center = T, scale = T)))%>%
  column_to_rownames(var = "country")
```

# PCA
```{r}
pca_correlates <- prcomp(poo)
summary(pca_correlates)

#10 PCs explain 92% of variation
pca_correlates$rotation[,1:10]

correlates_transform = as.data.frame(-pca_correlates$x[,1:10])

factoextra::fviz_nbclust(correlates_transform, kmeans, nstart = 100, method = 'wss')
factoextra::fviz_nbclust(correlates_transform, kmeans, method = 'silhouette')
factoextra::fviz_nbclust(correlates_transform, kmeans, method = 'gap_stat')
```

```{r}
pca_data <- pca_correlates$x %>%
  as.data.frame()%>%
  rownames_to_column()
full_data <- left_join(data.mean, pca_data, by = c("country" = "rowname"))
```

#Per country
```{r}
#set graphical parameters
par(mfrow=c(1,1), mar=c(3,2,0.1,1), oma=c(1,1,1,0), bg="white", fg="black") 

#PC1 and PC2
plot((full_data$PC1), (full_data$PC2), pch=21, bg=as.factor(full_data$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,1]*4),(pca_correlates$rotation[,2]*4), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,1]*4.1, pca_correlates$rotation[,2]*4.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((full_data$PC1), (full_data$PC2), data.mean$country, cex=0.5)

#PC2 and PC3
plot((full_data$PC2), (full_data$PC3), pch=21, bg=as.factor(full_data$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,2]*4),(pca_correlates$rotation[,3]*4), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,2]*4.1, pca_correlates$rotation[,3]*4.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((full_data$PC2), (full_data$PC3), data.mean$country, cex=0.5)

#PC3 and PC4
plot((full_data$PC3), (full_data$PC4), pch=21, bg=as.factor(full_data$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,3]*4),(pca_correlates$rotation[,4]*4), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,3]*4.1, pca_correlates$rotation[,4]*4.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((full_data$PC3), (full_data$PC4), data.mean$country, cex=0.5)

#PC4 and PC5
plot((full_data$PC4), (full_data$PC5), pch=21, bg=as.factor(full_data$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,4]*4),(pca_correlates$rotation[,5]*4), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,4]*4.1, pca_correlates$rotation[,5]*4.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((full_data$PC4), (full_data$PC5), data.mean$country, cex=0.5)
```

# Per region
```{r}
pca_region <- full_data %>% 
  group_by(iucn_region) %>% 
  summarise(mean_PC1 = mean(PC1), mean_PC2 = mean(PC2),mean_PC3 = mean(PC3), mean_PC4 = mean(PC4), mean_PC5 = mean(PC5),stdev1 = sd(PC1), stdev2 = sd(PC2), stdev3 = sd(PC3), stdev4 = sd(PC4), stdev5 = sd(PC5)) 
```

```{r}
#set graphical parameters
par(mfrow=c(1,1), mar=c(3,2,0.1,1), oma=c(1,1,1,0), bg="white", fg="black")

#PC1 and PC2
plot((pca_region$mean_PC1), (pca_region$mean_PC2), pch=21, bg=as.factor(pca_region$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
#lines((pca_region$mean_PC1 - stdev1), pca_region$mean_PC1, (pca_region$mean_PC1 + stdev1), pca_region$mean_PC1)
arrows(0,0,(pca_correlates$rotation[,1]*2),(pca_correlates$rotation[,2]*2), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,1]*2.1, pca_correlates$rotation[,2]*2.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((pca_region$mean_PC1), (pca_region$mean_PC2), pca_region$iucn_region, cex=0.5)

#PC2 and PC3
plot((pca_region$mean_PC2), (pca_region$mean_PC3), pch=21, bg=as.factor(pca_region$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,2]*1),(pca_correlates$rotation[,3]*1), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,2]*1.1, pca_correlates$rotation[,3]*1.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((pca_region$mean_PC2), (pca_region$mean_PC3), pca_region$iucn_region, cex=0.5)

#PC3 and PC4
plot((pca_region$mean_PC3), (pca_region$mean_PC4), pch=21, bg=as.factor(pca_region$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,3]*1),(pca_correlates$rotation[,4]*1), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,3]*1.1, pca_correlates$rotation[,4]*1.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((pca_region$mean_PC3), (pca_region$mean_PC4), pca_region$iucn_region, cex=0.5)

#PC4 and PC5
plot((pca_region$mean_PC4), (pca_region$mean_PC5), pch=21, bg=as.factor(pca_region$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,4]*1),(pca_correlates$rotation[,5]*1), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,4]*1.1, pca_correlates$rotation[,5]*1.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((pca_region$mean_PC4), (pca_region$mean_PC5), pca_region$iucn_region, cex=0.5)
```

# Models
```{r}
cons <- read.csv("Total_consumption_2012-2021.csv")

country_list <- unique(full_data$iso3_code)

full_data2 <- cons %>%
  filter(ISO3_Code %in% country_list) %>% 
  left_join(full_data, by = c("ISO3_Code" = "iso3_code")) 
```

```{r}
model1 <- glm(consumption~PC1, data = full_data2)
summary(model1)
```

Direct effect model
```{r}
model2 <- glm(consumption~mislabelling+culinary_traditions+religious_traditions+sensory_perception+health_beliefs+eating_habits+preparation+self_efficacy+price_perception+fish_availability+marketing+livelihood, data = full_data2)
summary(model2)
```

Upload DAG
```{r}
# install dagitty package
install.packages("dagitty")
library(dagitty)

# download specified DAG from dagitty.net 
DAG <- dagitty("dag {
culinary_traditions -> consumption
culinary_traditions -> eating_habits
eating_habits -> consumption
fish_availability -> consumption
fish_availability -> self_efficacy
health_beliefs -> consumption
livelihood -> consumption
marketing -> consumption
marketing -> eating_habits
marketing -> health_beliefs
measurement_error -> consumption
measurement_process -> measurement_error
mislabelling -> consumption
preparation -> consumption
preparation -> self_efficacy
price_perception -> consumption
religious_traditions -> consumption
religious_traditions -> culinary_traditions
self_efficacy -> consumption
sensory_perception -> consumption
}")

# plot DAG
plot(graphLayout(DAG))
```

Diagnostic tests
```{r}
# Creat new dataframe with matching columns
full_data3 <- full_data2 %>% 
  select(consumption, mislabelling:livelihood)
glimpse(full_data3)

# evaluate the d-separation implications of our DAG with our simulated dataset 
test <- localTests(DAG, full_data3)

# perform Holm-Bonferrino correction to mitigate problems around multiple testing 
test$p.value <- p.adjust(test$p.value) 

test # should show all p values above 0.05 and estimate below 0.3, suggesting DAG-data consistency
```

Setting flat priors
```{r}
bprior <- c(set_prior("normal(0,10)", dpar=c("self_efficacy", "fish_availability", "preparation")), prior_(~cauchy(0,2), class = ~Intercept), prior_(~cauchy(0,2), class = ~sigma))

#priors.cons <- c(prior(normal(0, 10), class = "b"))#, prior(normal(0, 50),coef =  ~Intercept)), prior(student_t(3, 0, 20), "sd"), prior(student_t(3, 0, 20), "sigma"))

#get_prior(consumption ~ self_efficacy + fish_availability + preparation, data = full_data3, family = gaussian())
```

Model (self-efficacy)
```{r}
model_slf <- brm(consumption ~ self_efficacy + fish_availability + preparation, data = full_data3, family = gaussian(), iter = 10000, save_pars(all), prior = bprior, control = list(adapt_delta = 0.99))
```

Model (mislabelling)
```{r}
model_mis <- brm(consumption ~ mislabelling, data = full_data3, family = gaussian(), iter = 10000, save_pars(all), prior = priors.cons, control = list(adapt_delta = 0.99))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

