---
title: "test"
author: "Ana Martins & Chris Mull"
date: "27/03/2023"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r load libraries, echo=FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(directlabels)
library(psych)
library(brms)
```

# Load data, as.factor 'country' and inspect
```{r clean consumption data, echo=FALSE, message=FALSE, eval=FALSE}
data <- read.csv("test1.csv") %>% 
  mutate(country = as.factor(country))
glimpse(data)
```

# Group, filter and summarise data
```{r}
data.new <- as.data.frame(apply(data,
                                2,
                                function(x) gsub("\\s+","",x)))
papers <- data.new %>% 
  group_by(country) %>% 
  filter(country!="All" & country!="All ") %>% 
  summarise(number_of_papers = n_distinct(id)) %>%
  mutate(error = 1/number_of_papers)
```

# Rank countries and IUCN regions by number of studies
```{r}
top10_country <- data.new %>% 
  group_by(country) %>%
  count() %>%
  arrange(desc(n))

top10_region <- data.new %>% 
  group_by(iucn_region) %>%
  count()%>%
  arrange(desc(n))
```

# Merge positive and negative correlates into one column
```{r}
correlates <- data.new %>%
  mutate(culinary_traditions = case_match(culinary_traditions_negative,
                                        "1" ~ "-1", 
                                        .default = culinary_traditions_positive),
         religious_traditions = case_match(religious_traditions_negative,
                                          "1" ~ "-1", 
                                        .default = religious_traditions_positive),
         sensory_perception = case_match(sensory_perception_negative,
                                         "1" ~ "-1", 
                                        .default = sensory_perception_positive),
         health_beliefs = case_match(health_beliefs_negative,
                                     "1" ~ "-1", 
                                        .default = health_beliefs_positive),
         eating_habits = case_match(eating_habits_negative,
                                    "1" ~ "-1", 
                                        .default = eating_habits_positive),
         preparation = case_match(preparation_negative,
                                  "1" ~ "-1", 
                                        .default = preparation_positive),
         self_efficacy = case_match(self_efficacy_negative,
                                    "1" ~ "-1", 
                                        .default = self_efficacy_positive),
         price_perception = case_match(price_perception_negative,
                                        "1" ~ "-1", 
                                        .default = price_perception_positive),
         fish_availability = case_match(fish_availability_negative,
                                         "1" ~ "-1", 
                                        .default = fish_availability_positive),
         marketing = case_match(marketing_negative,
                                 "1" ~ "-1", 
                                        .default = marketing_postive),
         livelihood = case_match(livelihood_negative,
                                 "1" ~ "-1", 
                                        .default = livelihood_positive))
```

# Select merged columns and make correlates numeric
```{r}
df <- correlates %>%
  select(iucn_region,
         country,
         iso3_code,
         mislabelling:livelihood)%>%
  mutate(across(mislabelling:livelihood, as.numeric))
str(df)
```

# Take the mean of each correlate 
```{r}
data.mean <- df %>% 
  group_by(country) %>%
  summarise(iucn_region = first(iucn_region),
            iso3_code = first(iso3_code),
            mislabelling = mean (mislabelling),
            culinary_traditions = mean (culinary_traditions),
            religious_traditions = mean (religious_traditions),
            sensory_perception = mean(sensory_perception),
            health_beliefs = mean(health_beliefs),
            eating_habits = mean(eating_habits),
            preparation = mean(preparation),
            self_efficacy = mean(self_efficacy),
            price_perception = mean (price_perception), 
            fish_availability = mean(fish_availability), 
            marketing = mean (marketing), 
            livelihood = mean (livelihood))
```

# Summarise columns and scale/center correlates
```{r}
poo <- data.mean %>% 
  select(-c(iucn_region,iso3_code))%>%
  mutate(across(mislabelling:livelihood, ~scale(., center = T, scale = T)))%>%
  column_to_rownames(var = "country")
```

# Run PCA analysis
```{r}
pca_correlates <- prcomp(poo)
summary(pca_correlates)

#10 PCs explain 92% of variation
pca_correlates$rotation[,1:10]

correlates_transform = as.data.frame(-pca_correlates$x[,1:10])

factoextra::fviz_nbclust(correlates_transform, kmeans, nstart = 100, method = 'wss')
factoextra::fviz_nbclust(correlates_transform, kmeans, method = 'silhouette')
factoextra::fviz_nbclust(correlates_transform, kmeans, method = 'gap_stat')
```

# Combine correlates and PC values in a single dataframe
```{r}
pca_data <- pca_correlates$x %>%
  as.data.frame()%>%
  rownames_to_column()

full_data <- left_join(data.mean, pca_data, by = c("country" = "rowname"))
```

# Plot PC values per country
```{r}
#set graphical parameters
par(mfrow=c(1,1), mar=c(3,2,0.1,1), oma=c(1,1,1,0), bg="white", fg="black") 

#PC1 and PC2
plot((full_data$PC1), (full_data$PC2), pch=21, bg=as.factor(full_data$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,1]*4),(pca_correlates$rotation[,2]*4), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,1]*4.1, pca_correlates$rotation[,2]*4.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((full_data$PC1), (full_data$PC2), data.mean$country, cex=0.5)

#PC2 and PC3
plot((full_data$PC2), (full_data$PC3), pch=21, bg=as.factor(full_data$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,2]*4),(pca_correlates$rotation[,3]*4), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,2]*4.1, pca_correlates$rotation[,3]*4.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((full_data$PC2), (full_data$PC3), data.mean$country, cex=0.5)

#PC3 and PC4
plot((full_data$PC3), (full_data$PC4), pch=21, bg=as.factor(full_data$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,3]*4),(pca_correlates$rotation[,4]*4), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,3]*4.1, pca_correlates$rotation[,4]*4.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((full_data$PC3), (full_data$PC4), data.mean$country, cex=0.5)

#PC4 and PC5
plot((full_data$PC4), (full_data$PC5), pch=21, bg=as.factor(full_data$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,4]*4),(pca_correlates$rotation[,5]*4), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,4]*4.1, pca_correlates$rotation[,5]*4.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((full_data$PC4), (full_data$PC5), data.mean$country, cex=0.5)
```

# Filter PC values per region and calculate the mean
```{r}
pca_region <- full_data %>% 
  group_by(iucn_region) %>% 
  summarise(mean_PC1 = mean(PC1), mean_PC2 = mean(PC2),mean_PC3 = mean(PC3), mean_PC4 = mean(PC4), mean_PC5 = mean(PC5),stdev1 = sd(PC1), stdev2 = sd(PC2), stdev3 = sd(PC3), stdev4 = sd(PC4), stdev5 = sd(PC5)) 
```

# Plot PC values per region
```{r}
#set graphical parameters
par(mfrow=c(1,1), mar=c(3,2,0.1,1), oma=c(1,1,1,0), bg="white", fg="black")

#PC1 and PC2
plot((pca_region$mean_PC1), (pca_region$mean_PC2), pch=21, bg=as.factor(pca_region$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
#lines((pca_region$mean_PC1 - stdev1), pca_region$mean_PC1, (pca_region$mean_PC1 + stdev1), pca_region$mean_PC1)
arrows(0,0,(pca_correlates$rotation[,1]*2),(pca_correlates$rotation[,2]*2), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,1]*2.1, pca_correlates$rotation[,2]*2.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((pca_region$mean_PC1), (pca_region$mean_PC2), pca_region$iucn_region, cex=0.5)

#PC2 and PC3
plot((pca_region$mean_PC2), (pca_region$mean_PC3), pch=21, bg=as.factor(pca_region$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,2]*1),(pca_correlates$rotation[,3]*1), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,2]*1.1, pca_correlates$rotation[,3]*1.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((pca_region$mean_PC2), (pca_region$mean_PC3), pca_region$iucn_region, cex=0.5)

#PC3 and PC4
plot((pca_region$mean_PC3), (pca_region$mean_PC4), pch=21, bg=as.factor(pca_region$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,3]*1),(pca_correlates$rotation[,4]*1), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,3]*1.1, pca_correlates$rotation[,4]*1.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((pca_region$mean_PC3), (pca_region$mean_PC4), pca_region$iucn_region, cex=0.5)

#PC4 and PC5
plot((pca_region$mean_PC4), (pca_region$mean_PC5), pch=21, bg=as.factor(pca_region$iucn_region), xlab="", ylab="", cex=1.5) #type = "null")
box(col="black", lwd=1.5)
arrows(0,0,(pca_correlates$rotation[,4]*1),(pca_correlates$rotation[,5]*1), length=0.1, angle=15, col="red", lwd=2.5)
text(pca_correlates$rotation[,4]*1.1, pca_correlates$rotation[,5]*1.1, label=c("Mislabelling","Culinary Traditions","Religious Traditions","Sensory Perception","Health Beliefs","Eating Habits","Preparation","Self Efficacy","Price Perception","Fish Availability","Marketing","Livelihood"))
abline(0,0, col="gray50", lty=2)
abline(v=0, col="gray50", lty=2)
text((pca_region$mean_PC4), (pca_region$mean_PC5), pca_region$iucn_region, cex=0.5)
```

## Bayesian Models
# Load data and filter by country and IUCN region
```{r}
cons <- read.csv("Total_consumption_2012-2021.csv")

country_list <- unique(full_data$iso3_code)

full_data2 <- cons %>%
  filter(ISO3_Code %in% country_list) %>% 
  left_join(full_data, by = c("ISO3_Code" = "iso3_code")) 
```

# Upload DAG from dagitty and plot it
```{r}
# install dagitty package
install.packages("dagitty")
library(dagitty)

# download specified DAG from dagitty.net 
DAG <- dagitty("dag {
culinary_traditions -> consumption
culinary_traditions -> eating_habits
eating_habits -> consumption
fish_availability -> consumption
fish_availability -> self_efficacy
health_beliefs -> consumption
livelihood -> consumption
marketing -> consumption
marketing -> eating_habits
marketing -> health_beliefs
measurement_error -> consumption
measurement_process -> measurement_error
mislabelling -> consumption
preparation -> consumption
preparation -> self_efficacy
price_perception -> consumption
religious_traditions -> consumption
religious_traditions -> culinary_traditions
self_efficacy -> consumption
sensory_perception -> consumption
}")

# plot DAG
plot(graphLayout(DAG))
```

# Run diagnostic tests
```{r}
# Creat new dataframe with matching columns
full_data3 <- full_data2 %>% 
  left_join(papers, by = "country")%>%
  select(country, consumption, mislabelling:livelihood, error) 
  
glimpse(full_data3)

# Evaluate the d-separation implications of our DAG with our simulated dataset 
test <- localTests(DAG, full_data3)

# Perform Holm-Bonferrino correction to mitigate problems around multiple testing 
test$p.value <- p.adjust(test$p.value) 

# test should show all p values above 0.05 and estimate below 0.3, suggesting DAG-data consistency
```

# Standirdise consumption estimates
```{r}
full_data4 <- full_data3 %>% 
  mutate(consumption = scale(consumption, center = T, scale = T))
```

# Model A (mislabelling)
```{r}
# Set flat priors for all models
aprior <- prior(normal(0,10), class = b, coef = mislabelling)# No open biasing paths

model_mis <- brm(consumption ~ mislabelling, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = aprior, control = list(adapt_delta = 0.99))
summary(model_mis)
plot(model_mis)
```

# Model B (culinary_tradition)
```{r}
bprior <- c(prior(normal(0,10), class = b, coef = culinary_traditions), prior(normal(0,10), class = b, coef = religious_traditions)) # Biasing paths are open (religious_traditions)

model_cul <- brm(consumption ~ culinary_traditions + religious_traditions, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = bprior, control = list(adapt_delta = 0.99))

summary(model_cul)
plot(model_cul)
```

# Model C (religious_traditions)
```{r}
cprior <- prior(normal(0,10), class = b, coef = religious_traditions)# No open biasing paths

model_rel <- brm(consumption ~ religious_traditions, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = cprior, control = list(adapt_delta = 0.99))
summary(model_rel)
plot(model_rel)
```

# Model D (sensory_perception)
```{r}
dprior <- prior(normal(0,10), class = b, coef = sensory_perception)# No open biasing paths

model_sen <- brm(consumption ~ sensory_perception, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = dprior, control = list(adapt_delta = 0.99))
summary(model_sen)
plot(model_sen)
```

# Model E (health_beliefs)
```{r}
eprior <- c(prior(normal(0,10), class = b, coef = health_beliefs), prior(normal(0,10), class = b, coef = marketing)) # Biasing paths are open (marketing).

model_hea <- brm(consumption ~ health_beliefs + marketing, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = eprior, control = list(adapt_delta = 0.99))

summary(model_hea)
plot(model_hea)
```

# Model F (eating_habits)
```{r}
fprior <- c(prior(normal(0,10), class = b, coef = culinary_traditions), prior(normal(0,10), class = b, coef = marketing), prior(normal(0,10), class = b, coef = eating_habits)) # Biasing paths are open (culinary_traditions, marketing).

model_eat <- brm(consumption ~ eating_habits + culinary_traditions + marketing, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = fprior, control = list(adapt_delta = 0.99))

summary(model_eat)
plot(model_eat)
```

# Model G (preparation)
```{r}
gprior <- prior(normal(0,10), class = b, coef = preparation)# No open biasing paths

model_pre <- brm(consumption ~ preparation, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = gprior, control = list(adapt_delta = 0.99))
summary(model_pre)
plot(model_pre)
```

# Model H (self_efficacy)
```{r}
hprior <- c(prior(normal(0,10), class = b, coef = fish_availability), prior(normal(0,10), class = b, coef = preparation), prior(normal(0,10), class = b, coef = self_efficacy)) # Biasing paths are open (fish_availability, preparation).

model_sel <- brm(consumption ~ self_efficacy + fish_availability + preparation, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = hprior, control = list(adapt_delta = 0.99))

summary(model_sel)
plot(model_sel)
```

# Model I (price_perception)
```{r}
iprior <- prior(normal(0,10), class = b, coef = price_perception)# No open biasing paths

model_pri <- brm(consumption ~ price_perception, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = iprior, control = list(adapt_delta = 0.99))
summary(model_pri)
plot(model_pri)
```

# Model J (fish_availability)
```{r}
jprior <- prior(normal(0,10), class = b, coef = fish_availability)# No open biasing paths

model_fis <- brm(consumption ~ fish_availability, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = jprior, control = list(adapt_delta = 0.99))
summary(model_fis)
plot(model_fis)
```

# Model K (marketing)
```{r}
kprior <- prior(normal(0,10), class = b, coef = marketing)# No open biasing paths

model_mar <- brm(consumption ~ marketing, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = kprior, control = list(adapt_delta = 0.99))
summary(model_mar)
plot(model_mar)
```

# Model L (livelihood)
```{r}
lprior <- prior(normal(0,10), class = b, coef = livelihood)# No open biasing paths

model_liv <- brm(consumption ~ livelihood, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), prior = lprior, control = list(adapt_delta = 0.99))
summary(model_liv)
plot(model_liv)
```

# Model M (measurament error)
```{r}
#prior <- c(prior(normal(0,10), class = b, coef = culinary_traditions), prior(normal(0,10), class = b, coef = marketing), prior(normal(0,10), class = b, coef = eating_habits))

model_err <- brm(consumption ~ error, data = full_data4, family = gaussian(), iter = 10000, save_pars = save_pars(all = TRUE), control = list(adapt_delta = 0.99))

summary(model_err)
plot(model_err)
```

# Map of total global consumption
```{r}
library(rnaturalearth)
library(rnaturalearthdata)
library(ggplot2)
library(sf)
library(tidyverse)
library(sf)
library(viridis)

# Load data
total_consumption <- read.csv("Total_consumption_2012-2021.csv")

world <- ne_countries(scale = "medium", returnclass = "sf")

# Linking data frames
data1 <- total_consumption %>%
  left_join(world, by = c("ISO3_Code" = "brk_a3"))

# Mapping
ggplot(data = data1) +
    geom_sf(aes(fill = consumption, geometry = geometry)) +
            xlab("Longitude") + ylab("Latitude") +
            ggtitle("Global Consumprtion of Shark and Ray Meat") +
   scale_fill_viridis_c(option = "viridis", trans = "sqrt", name="Total consumption (mt)", na.value = "grey50")
```

# Map of global consumption per capita
```{r}
# Load data
pop100km <- read.csv("nagdc-population-landscape-and-climate-estimates-version-3.csv")

# Linking data frames
data2 <- pop100km %>%
  filter(THEMENAME=="COASTAL PROXIMITY: 100KM ZONE") %>% 
  left_join(data1, by = c("ISO3V10" = "ISO3_Code")) 

# Calculate per capita consumption
as.numeric(data2$POPULATION2010)
data3 <- mutate(data2, consumption_pc = consumption/POPULATION2010)
plot(data3$consumption_pc)

# Mapping
 ggplot(data = data3) +
    geom_sf(aes(fill = consumption_pc, geometry = geometry)) +
            xlab("Longitude") + ylab("Latitude") +
            ggtitle("Global Consumprtion of Shark and Ray Meat") +
   scale_fill_viridis_c(option = "viridis", trans = "sqrt", name="Per capita consumption", na.value = "grey50", breaks=c(0, 0.02,0.04,0.06,0.08, 1)) +
  theme(panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.background = element_rect(fill = 'white', colour = 'white'))
```

